Negative Perceptions of Outsourcing to Artificial Intelligence

### 2025-05-01 manuscript.qmd
Further edits for submission

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: Word count (excl. Methods and Results) = 4619 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: >
  As artificial intelligence (AI) tools become increasingly integrated into 
  daily life, people are beginning to outsource not only professional tasks but 
  also socio-relational ones. Large language models like ChatGPT can generate 
  wedding vows, speeches, and personal messages, raising questions about how 
  individuals who use AI for such tasks are perceived by others. In this paper,
  we conduct five pre-registered studies with British participants (N = 3,649)
  to understand how people view those who outsource tasks to AI, and how this
  depends on how socio-relational the task is, whether AI is used as a tool or
  fully delegated to, and the acknowledgment of the AI use. We find negative
  perceptions of outsourcing, particularly for socio-relational tasks. We show
  that outsourcing makes us think more negatively about not only the person and
  their motivations, but also the outsourced work itself. Moreover, we provide
  insight into why this occurs: the reduced effort from outsourcing
  socio-relational tasks to AI signals that the output is less authentically 
  one's own and that the person cares less about the task. Our research 
  highlights the way that AI use shapes our perceptions of people, raising key
  philosophical questions about efficiency, authenticity, and social ties in a
  world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-pdf: default
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

@@

## Present Research

@@

# Study 1

## Methods

### Ethical Approval

@@

### Participants

@@

### Design

@@

### Procedure

@@

### Pre-registration

@@

### Statistical Analysis

@@

### Transparency and Openness

@@

## Results

@@

## Discussion

@@

# Study 2

## Methods

### Participants

@@

# Study 3

## Methods

### Participants

@@

# Study 4

## Methods

### Participants

@@

# Study 5

## Methods

### Participants

@@

# General Discussion

@@

## Limitations and Directions for Future Research

@@

## Conclusions

@@

# Acknowledgements

@@

# Data and Code Availability

@@

# Statement of Interests

@@

# References

>>>>

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: Word count (excl. Methods and Results) = 4619 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: >
  As artificial intelligence (AI) tools become increasingly integrated into 
  daily life, people are beginning to outsource not only professional tasks but 
  also socio-relational ones. Large language models like ChatGPT can generate 
  wedding vows, speeches, and personal messages, raising questions about how 
  individuals who use AI for such tasks are perceived by others. In this paper,
  we conduct five pre-registered studies with British participants (N = 3,649)
  to understand how people view those who outsource tasks to AI, and how this
  depends on how socio-relational the task is, whether AI is used as a tool or
  fully delegated to, and the acknowledgment of the AI use. We find negative
  perceptions of outsourcing, particularly for socio-relational tasks. We show
  that outsourcing makes us think more negatively about not only the person and
  their motivations, but also the outsourced work itself. Moreover, we provide
  insight into why this occurs: the reduced effort from outsourcing
  socio-relational tasks to AI signals that the output is less authentically 
  one's own and that the person cares less about the task. Our research 
  highlights the way that AI use shapes our perceptions of people, raising key
  philosophical questions about efficiency, authenticity, and social ties in a
  world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-pdf:
    numbered-lines: true
    include-in-header:
      text: \nolinenumbers
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

\linenumbers

@@

## Present Research {.unnumbered .unlisted}

@@

# Study 1 {.unnumbered .unlisted}

## Methods {.unnumbered .unlisted}

### Ethical Approval {.unnumbered .unlisted}

@@

### Participants {.unnumbered .unlisted}

@@

### Design {.unnumbered .unlisted}

@@

### Procedure {.unnumbered .unlisted}

@@

### Pre-registration {.unnumbered .unlisted}

@@

### Statistical Analysis {.unnumbered .unlisted}

@@

### Transparency and Openness {.unnumbered .unlisted}

@@

## Results {.unnumbered .unlisted}

@@

## Discussion {.unnumbered .unlisted}

@@

# Study 2 {.unnumbered .unlisted}

## Methods {.unnumbered .unlisted}

### Participants {.unnumbered .unlisted}

@@

# Study 3 {.unnumbered .unlisted}

## Methods {.unnumbered .unlisted}

### Participants {.unnumbered .unlisted}

@@

# Study 4 {.unnumbered .unlisted}

## Methods {.unnumbered .unlisted}

### Participants {.unnumbered .unlisted}

@@

# Study 5 {.unnumbered .unlisted}

## Methods {.unnumbered .unlisted}

### Participants {.unnumbered .unlisted}

@@

# General Discussion {.unnumbered .unlisted}

@@

## Limitations and Directions for Future Research {.unnumbered .unlisted}

@@

## Conclusions {.unnumbered .unlisted}

@@

\newpage
\nolinenumbers

# Acknowledgements {.unnumbered .unlisted}

@@

# Data and Code Availability {.unnumbered .unlisted}

@@

# Statement of Interests {.unnumbered .unlisted}

@@

# References {.unnumbered .unlisted}
